Notes on filtration of logics containing {\bf K5}

### 2021-02-10 filtration-stuff/Text.tex
fix finite 'canonical' stuff

\section{``Finite'' canonical models}

@@

\begin{defin}
  Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$. A ``finite'' canonical model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
  \begin{enumerate}
    \item $W_{\mathcal{L}}^{\varphi}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
    \item $\Delta_1 R \Delta_2$ iff $\Diamond \psi \in \Delta_2 \Rightarrow \psi \in \Delta_1$
    \item $\vartheta_{\mathcal{L}}^{\varphi}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$.
  \end{enumerate}
\end{defin}

@@

\begin{theorem} Let $\mathcal{L}$ be a complete normal modal logic, then the following are equivalent:
  \begin{enumerate}
    \item $\mathcal{L}$ admits strict filtrations
    \item If $\varphi \not\in \mathcal{L}$, then the ``finite'' canonical model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
  \end{enumerate}
\end{theorem}

@@

    Consider a ``finite'' canonical model
    $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} = \langle W_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, R_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, \vartheta_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \rangle$.
    Let us show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}}$ refutes $\varphi$. For that, we show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$.

@@

    Let $\mathcal{M}_{\mathcal{L}}^{\varphi} = $ be a ``finite'' canonical model that refutes $\varphi$.

@@

    $R_{\mathcal{L}}^{\varphi}$ also trivially satisfies the required defined. Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite. In particular, ${\bf K}5$ does not have the ``finite'' canonical property that follows from the contraposition of this statement and Theorem~\ref{sad}.

>>>>

\section{Finite ``canonical`` models}

@@

\begin{defin}
  Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$. A finite ``canonical'' model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
  \begin{enumerate}
    \item $W_{\mathcal{L}}^{\varphi}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
    \item $R_{\mathcal{L}}^{\varphi}$ is a relation such that
    $\langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi} \rangle$ is an $\mathcal{L}$-frame
    and
    \begin{center}
    $\forall \Box \psi \operatorname{Sub}(\varphi) \:\:
    \forall \Delta_1 \in W_{\mathcal{L}}^{\varphi} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
    \forall \Delta_2 \in R_{\mathcal{L}}^{\varphi}(\Delta_1) \:\: \psi \in \Delta_2)$
    \end{center}
    \item $\vartheta_{\mathcal{L}}^{\varphi}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$.
  \end{enumerate}
\end{defin}

@@

\begin{theorem} Let $\mathcal{L}$ be a complete normal modal logic, then the following are equivalent:
  \begin{enumerate}
    \item $\mathcal{L}$ admits strict filtrations
    \item If $\varphi \not\in \mathcal{L}$, then the finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
  \end{enumerate}
\end{theorem}

@@

    Consider a finite ``canonical'' model
    $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} = \langle W_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, R_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, \vartheta_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \rangle$.
    Let us show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}}$ refutes $\varphi$. For that, we show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$.

@@

    Let $\mathcal{M}_{\mathcal{L}}^{\varphi}$ be a finite ``canonical'' model that refutes $\varphi$.

@@

    $R_{\mathcal{L}}^{\varphi}$ also satisfies the required definition.
    Suppose $\Box \psi \in \varphi$ and
    $\mathcal{M}_{\mathcal{L}}^{\varphi}, \Delta_1 \models \Box \psi$ and
    $\Delta_1 R_{\mathcal{L}}^{\varphi} \Delta_2$. Clearly that $\psi \in \Delta_2$ by the definition
    of a finite ``canonical'' model.

    Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite. In particular, ${\bf K}5$ fails to have the ``canonical'' model property that follows from the contraposition of this statement and Theorem~\ref{sad}.
